---
title: "Case Study"
output:
  html_document: 
    toc: true
    toc_float: true
---

Pulling things together.

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

library(tidyverse)
```

## Some slides

<script async class="speakerdeck-embed" data-id="69791e0a56b14efd9f741fc10cea30f9" data-ratio="1.33333333333333" src="//speakerdeck.com/assets/embed.js"></script>
<div style="margin-bottom:5px"> <strong> <a href="https://speakerdeck.com/jeffgoldsmith/dsi-data-science-revisited" title="'Data Science' Revisited" target="_blank">"Data Science" Revisited</a> </strong> from <strong><a href="https://speakerdeck.com/jeffgoldsmith" target="_blank">Jeff Goldsmith</a></strong>. </div><br>


## Example

This example is based entirely on live-coding, and uses the [Instacart dataset](dataset_instacart.html). These data can be loaded directly from the course website using:

```{r eval = FALSE}
instacart_data = read_csv("./data/instacart_train_data.csv.zip")
```

After some quick exploratory analyses, we'll brainstorm some questions and try to answer them. 



## Other materials

* If you didn't listen before, now's the time for Chris Volinsky's ["How Industry Views Data Science Education in Statistics Departments"](https://youtu.be/JLs01Z5baSU?t=1063)
* Jeff Leek's [problem forward](https://simplystatistics.org/2013/05/29/what-statistics-should-do-about-big-data-problem-forward-not-solution-backward/) blog post
