---
title: "Exploratory analysis using `group_by`"
output:
  html_document: 
    toc: true
    toc_float: true
---

Data sets can frequently be partitioned into meaningful groups based on the variables they contain. Making this grouping explicit allows the computation of numeric summaries within groups, which in turn facilitates quantitative comparisons. 

This is the second module in the [Visualization and EDA](topic_visualization_and_eda.html) topic; the relevant slack channel is [here](https://p8105-fall2017.slack.com/messages/C7CGD92CX).

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

library(tidyverse)
options(tibble.print_min = 6)
```

## Some slides

<script async class="speakerdeck-embed" data-id="38db98aa225d41a0a0d7fec84ae7a259" data-ratio="1.33333333333333" src="//speakerdeck.com/assets/embed.js"></script>
<div style="margin-bottom:5px"> <strong> <a href="https://speakerdeck.com/jeffgoldsmith/dsi-exploratory-analysis" title="Exploratory analysis" target="_blank">Exploratory Analysis</a> </strong> from <strong><a href="https://speakerdeck.com/jeffgoldsmith" target="_blank">Jeff Goldsmith</a></strong>. </div><br>


## Example

We'll continue in the same project that we used for visualization, and use essentially the same dataset -- the only exception is the addition of `month` variable, created using `lubridate::floor_date()`. 

```{r weather_data_create, cache = TRUE}
library(rnoaa)

weather = 
  meteo_pull_monitors(c("USW00094728", "USC00519397", "USS0023B17S"),
                      var = c("PRCP", "TMIN", "TMAX"), 
                      date_min = "2016-01-01",
                      date_max = "2016-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY", 
                      USC00519397 = "Waikiki_HA",
                      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10,
    month = lubridate::floor_date(date, unit = "month")) %>%
  select(name, id, date, month, everything())
weather
```

### `group_by`

Datasets are often comprised of groups defined by one or more (categorical) variable; `group_by()` makes these groupings explicit so that they can be included in subsequent operations. For example, we might group `weather` by `name` and `month`:

```{r}
weather %>%
  group_by(name, month)
```

Several important functions respect grouping structures. You will frequently use `summarize` to create one-number summaries within each group, or use `mutate` to define variables within groups. The rest of this example shows these functions in action.

Because these (and other) functions will use grouping information if it exists, it is sometimes necessary to remove groups using `ungroup()`.

### Counting things

As an intro to `summarize`, let's count the number of observations in each month in the complete `weather` dataset.

```{r}
weather %>%
  group_by(month) %>%
  summarize(n = n())
```

The result is a dataframe that includes the grouping variable and the desired summary. 

In this case, you could use `count()` in place of `group_by()` and `summarize()` if you remember that this function exists.

```{r}
weather %>%
  count(month)
```

You can use `summarize()` to compute multiple summaries within each group. As an example, we count the number of observations in each month and the number of distinct values of `date` in each month.

```{r}
weather %>%
  group_by(month) %>%
  summarize(n_obs = n(),
            n_days = n_distinct(date))
```


### General summaries

Standard statistical summaries are regularly computed in `summarize()` using functions like `mean()`, `median()`, `var()`, `sd()`, `mad()`, `IQR()`, `min()`, and `max()`. To use these, you indicate the variable to which they apply and include any additional arguments as necessary. 

```{r}
weather %>%
  group_by(month) %>%
  summarize(mean_tmax = mean(tmax),
            mean_prec = mean(prcp, na.rm = TRUE),
            median_tmax = median(tmax),
            sd_tmax = sd(tmax))
```

You can group by more than one variable. 

```{r}
weather %>%
  group_by(name, month) %>%
  summarize(mean_tmax = mean(tmax),
            median_tmax = median(tmax))
```

The fact that `summarize()` produces a dataframe is important (and consistent with other functions in the `tidyverse`). You can incorporate grouping and summarizing within broader analysis pipelines. For example, we can take create a plot based on the monthly summary: 

```{r}
weather %>%
  group_by(name, month) %>%
  summarize(mean_tmax = mean(tmax)) %>%
  ggplot(aes(x = month, y = mean_tmax, color = name)) + 
    geom_point() + geom_line() + 
    theme(legend.position = "bottom")
```

### Grouped `mutate`

Summarizing collapses groups into single data points. In contrast, using `mutate()` in conjuntion with `group_by()` will retain all original data points and add new variables computed within groups.

Suppose you want to compare the daily max temperature to the annual average max temperature for each station separately, and to plot the result. You could do so using:

```{r}
weather %>%
  group_by(name) %>%
  mutate(centered_tmax = tmax - mean(tmax)) %>% 
  ggplot(aes(x = date, y = centered_tmax, color = name)) + 
    geom_point() 
```


### Window functions

The previous example used `mean()` to compute the mean within each group, which was then subtracted from the observed max tempurature. `mean()` takes `n` inputs and produces a single output. 

Window functions, in contrast, take `n` inputs and return `n` outputs, and the outputs depend on all the inputs. There are several categories of window functions; you're most likely to need ranking functions and offsets, which we illustrate below.

First, we can find the max temperature ranking within month.
```{r}
weather %>%
  group_by(name, month) %>%
  mutate(temp_ranking = min_rank(tmax))
```

This sort of ranking is useful when filtering data based on rank. We could, for example, keep only the day with the lowest max temperature within each month:
```{r}
weather %>%
  group_by(name, month) %>%
  filter(min_rank(tmax) < 2)
```

We could also keep the three days with the highest max temperature:
```{r}
weather %>%
  group_by(name, month) %>%
  filter(min_rank(desc(tmax)) < 4)
```

In both of these, we've skipped a `mutate()` statement that would create a ranking variable, and gone straight to filtering based on the result. 

Offsets, especially lags, are used to compare an observation to it's previous value. This is useful, for example, to find the day-by-day change in max temperature within each station over the year:

```{r}
weather %>%
  group_by(name) %>%
  mutate(temp_change = tmax - lag(tmax))
```

This kind of variable might be used to quantify the day-by-day variability in max temperature, or to identify the largest one-day increase:

```{r}
weather %>%
  group_by(name) %>%
  mutate(temp_change = tmax - lag(tmax)) %>%
  summarize(temp_change_sd = sd(temp_change, na.rm = TRUE),
            temp_change_max = max(temp_change, na.rm = TRUE))
```


### Limitations

`summarize()` can only be used with functions that return a single-number summary. This creates a ceiling, even if it is **very** high. Later we'll see how to aggregate data in a more general way, and how to perform complex operations on the resulting sub-datasets.

## Other materials

* The preceding is based heavily on Jenny Bryan's [group_by](http://stat545.com/block010_dplyr-end-single-table.html#group_by-is-a-mighty-weapon) material
* R for Data Science has sub-chapters on [grouped summaries](http://r4ds.had.co.nz/transform.html#grouped-summaries-with-summarise) and [exploratory data analysis](http://r4ds.had.co.nz/exploratory-data-analysis.html)
* A more in-depth overview of window functions is [here](https://cran.r-project.org/web/packages/dplyr/vignettes/window-functions.html)

The code that I produced working examples in lecture is [here](./lecture_code/visualization_and_eda.zip).