---
title: "Interactive graphics and dashboards"
output:
  html_document: 
    toc: true
    toc_float: true
---

Static plots came up in [visualization](visualization.html); now we're going to see how you can interact with plots to increase the amount of information available. We'll also embed these in dashboards, which provide a way to give a complete picture of a dataset. 

This is the first module in the [Interactivity](topic_interactivity.html) topic; the relevant slack channel is [here](ZZZZZZZ).

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)
```


## Some slides

<script async class="speakerdeck-embed" data-id="bfb454627b71408889958f0893413097" data-ratio="1.33333333333333" src="//speakerdeck.com/assets/embed.js"></script>
<div style="margin-bottom:5px"> <strong> <a href="https://speakerdeck.com/jeffgoldsmith/dsi-strings-and-factors" title="Strings and Factors" target="_blank">Strings and Factors</a> </strong> from <strong><a href="https://speakerdeck.com/jeffgoldsmith" target="_blank">Jeff Goldsmith</a></strong>. </div><br>


## Example

Start directory, add r project, initialize git, add to client, put an R markdown file. 

```{r}
library(tidyverse)


theme_set(theme_bw())
theme_update(legend.position = "bottom")
```

### Plotly

Scatterplots, barcharts, boxplots

ggplotly

### Highcharts

Maybe for a treemap? 

### Dashboard

start a new r markdown file; add plots generated above in rows, columns, tabs. 

### Publishing a dashboard

create a webpage around this -- just use the website template, change the site nav and file names. publish on github (after changing GH settings).




## Other materials

* plotly: https://plot.ly/r/
* Highcharts: http://jkunst.com/highcharter/index.html
    * examples: http://jkunst.com/highcharter/showcase.html
* flexdashboards: http://rmarkdown.rstudio.com/flexdashboard/index.html
    * layouts: http://rmarkdown.rstudio.com/flexdashboard/layouts.html
    * examples: http://rmarkdown.rstudio.com/flexdashboard/examples.html

* pretty slick example: https://juliasilge.shinyapps.io/power-app/
    * motivation: https://juliasilge.com/blog/ab-testing/
  
The code that I produced working examples in lecture is [here]().
