---
title: "What is Data Science?"
output:
  html_document: 
    toc: true
    toc_float: true
---

We have, by now, established some fundamental tools for doing data science. It's important to revisit our definition, and especially our discussion of connotation, before moving forward. 

The slack channel for today's example is [here](https://p8105-fall2017.slack.com/messages/C7ET6NQGG).

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

library(tidyverse)
```

## Some slides

<script async class="speakerdeck-embed" data-id="69791e0a56b14efd9f741fc10cea30f9" data-ratio="1.33333333333333" src="//speakerdeck.com/assets/embed.js"></script>
<div style="margin-bottom:5px"> <strong> <a href="https://speakerdeck.com/jeffgoldsmith/dsi-data-science-revisited" title="'Data Science' Revisited" target="_blank">"Data Science" Revisited</a> </strong> from <strong><a href="https://speakerdeck.com/jeffgoldsmith" target="_blank">Jeff Goldsmith</a></strong>. </div><br>


## Example

This example is based entirely on live-coding and uses the [NYC Airbnb data](dataset_airbnb.html). Once downloaded, the data can be imported using:

```{r cache = TRUE}
airbnb_data = read_csv("./data/nyc_airbnb.zip")
```

As always, I'll do today's coding in a R Markdown file, sitting in an R Project with a data directory.

### Understanding variables

First, let's take a few minutes to understand the dataset and the variables it contains.

```{r}
# View(airbnb_data)
str(airbnb_data)

airbnb_data %>%
  count(room_type)

airbnb_data %>%
  count(neighbourhood_group)
```

### Brainstorming questions

A major element of data science is to ask questions, and this dataset provides some rich opportunities. For example, we might ask:

* Does rating vary by neighborhood, room type, or both?
* Do hosts with more listings have higher or lower ratings?
* How is price related to other variables?
* Where are rentals located?

We'll take a few minutes as a class to brainstorm some additional questions, and then try to answer some of them.

```{r, echo = FALSE, eval = FALSE}
ggplot(airbnb_data, aes(x = longitude, y = latitude)) +
  geom_point() + 
  coord_map()
  coord_equal(ratio = 1)
```


```{r, echo = FALSE, eval = FALSE}
library(sp)
library(rgdal)
library(rgeos)
library(ggplot2)
library(ggthemes)

fils <- unzip("~/Desktop/nybb_17c.zip")

nyc <- readOGR(fils[1], ogrListLayers(fils[1])[1], stringsAsFactors=FALSE)

nyc_map <- fortify(gSimplify(nyc, 0.05))

gg <- ggplot()
gg <- gg + geom_map(data=nyc_map, map=nyc_map,
                    aes(x=long, y=lat, map_id=id),
                    color="black", fill="white", size=0.25)
gg <- gg + coord_equal() 
gg <- gg + theme_map()
gg = gg + geom_point(data = airbnb_data, aes(x = longitude, y = latitude)) 
```

## Other materials

* Although we're playing with public (scraped) data, you might be interested in [How R Helps Airbnb Make the Most of Its Data](https://peerj.com/preprints/3182/)
* If you didn't listen before, now's the time for Chris Volinsky's ["How Industry Views Data Science Education in Statistics Departments"](https://youtu.be/JLs01Z5baSU?t=1063)
* Jeff Leek's [problem forward](https://simplystatistics.org/2013/05/29/what-statistics-should-do-about-big-data-problem-forward-not-solution-backward/) blog post

The code that I produced working examples in lecture is [here](./lecture_code/what_is_ds_ii.zip).